---
title: "Data_Clean_R"
author: "Zijing Cheng zc233"
date: "2023-04-22"

output: html_document
---

## Data Processing with R

```{r}
# load the tidyverse library
library(tidyverse)

# read the business.csv file into a dataframe
business = read.csv("../../data/clean_data/business.csv")

# use mutate to create a new column called "City" based on the "City" column,
# replacing any occurrences of "St. Petersburg" with "Saint Petersburg"
business = business %>% 
  mutate(City = ifelse(City == "St. Petersburg", "Saint Petersburg", City))

```

## Select cities with the top ten restaurants in each state
```{r}
# group business data by state and city, and summarize number of businesses
business_top = business %>% 
  group_by(State, City) %>% 
  summarise(n = n())

# group by state, calculate rank based on descending order of number of businesses, and filter top 10
business_top = business_top %>% 
  group_by(State) %>% 
  mutate(rank = rank(-n)) %>% 
  filter(rank <= 10)

# add a new column indicating whether business is open or closed based on 'is_open' column
business = business %>% 
  mutate(`Business Status` = ifelse(is_open == 0, "Close", "Open"))

```


## Write the CSV
```{r}
write.csv(business_top,"data/clean_data/business_top.csv")
write.csv(business,"data/clean_data/business.csv")
```

## Subset to the Florida dataset
```{r}
businessFL = business%>%filter(State == "FL")
businessFL = businessFL[,-c(1:4,18)]
summary(businessFL)
```


## Classified by restaurant region of categories
Here, we divided it into American, Asian, European, Latin American
```{r}
# Define a function to categorize businesses based on a list of categories
categorize_business <- function(data, category_list) {
  categorized_data <- data.frame()
  for (i in 1:nrow(data)) {
    row <- data[i,]
    if (sum(unlist(strsplit(row$categories, ", ")) %in% category_list) > 0) {
      row <- row %>% mutate(`Region Category` = category_list[[1]])
      categorized_data <- rbind(categorized_data, row)
    }
  }
  return(categorized_data)
}

# Categorize businesses based on predefined categories
businessFL1 = businessFL
categories = list(
  c('Southern','Barbeque','Fast Food','Diners','Dells','American(Traditional)','American(New)','Soup','Comfort Food','Breakfast & Brunch', 'Burgers','Hot Dogs','Chicken Wings','Chicken Shop','Fast Food',  'Sandwiches' ),
  c('Asian Fusion','Chinese','Japanese','Sushi Bars','Thai','Vietnamese'),
  c('Italian','Spanish','Steakhouses','Mediterranean'),
  c('Latin American','Cuban','Mexican','Tex-Mex','Caribbean')
)
businessFL_American = categorize_business(businessFL1, categories[[1]])
businessFL_Asian = categorize_business(businessFL1, categories[[2]])
businessFL_European = categorize_business(businessFL1, categories[[3]])
businessFL_Latin = categorize_business(businessFL1, categories[[4]])

# Combine categorized data into one dataframe
a = rbind(businessFL_Latin, businessFL_European)
a = rbind(a, businessFL_Asian)
a = rbind(a, businessFL_American)

# Remove duplicates based on business_id
a = a %>% group_by(business_id) %>% filter (!duplicated(business_id))

# Count number of businesses in each region category
a %>% group_by(`Region Category`) %>% summarize(Count = n())

```


## Classified by restaurant format of categories
Here, we divided it into Bars, Food Trucks, Food Delivery, Food Stores, Juice Bars, Desserts, Events
```{r}
# Function to categorize businesses based on a list of categories
# Input: data - dataframe with business data, category_list - list of categories to search for, category_name - name of category to assign
# Output: dataframe with businesses assigned to their respective category
categorize_business <- function(data, category_list, category_name) {
  categorized_data <- data.frame() # create an empty dataframe to hold categorized businesses
  for (i in 1:nrow(data)) { # loop through each row of the input data
    row <- data[i,] # get the current row
    if (sum(unlist(strsplit(row$categories, ", ")) %in% category_list) > 0) { # if any of the categories in the list are found in the current row, assign it to the specified category
      row <- row %>% mutate(`Format Category` = category_name) # add a new column `Format Category` to the row and assign it to the specified category
      categorized_data <- rbind(categorized_data, row) # add the categorized row to the dataframe
    }
  }
  return(categorized_data) # return the dataframe with categorized businesses
}

# Assign each business to its respective category using the categorize_business function and the list of categories
businessFL1 = businessFL # copy original dataframe
categories <- list(
  list('Beer','Beer Bar','Cocktall Bars', 'Pubs', 'Sports Bars', 'Lounges','Wine Bars'),
  list('Food Trucks'),
  list('Food Delivery Services'),
  list('Grocery','Specialty Food'),
  list('Juice Bars & Smoothies'),
  list('Ice Cream & Frozen Yogurt', 'Bakeries'),
  list('Music Venues','Event Planning & Services','Arts & Entertainment','Nightlife','Venues & Event Spaces','')
)

businessFL_Bars = categorize_business(businessFL1, categories[[1]], "Bars")
businessFL_trucks = categorize_business(businessFL1, categories[[2]], "Food Trucks")
businessFL_delivery = categorize_business(businessFL1, categories[[3]], "Food Delivery")
businessFL_store = categorize_business(businessFL1, categories[[4]], "Food Stores")
businessFL_Juice = categorize_business(businessFL1, categories[[5]], "Juice Bars")
businessFL_desserts = categorize_business(businessFL1, categories[[6]], "Desserts")
businessFL_events = categorize_business(businessFL1, categories[[7]], "Events")

# Combine all categorized dataframes into one
b = rbind(businessFL_Bars, businessFL_trucks, businessFL_delivery, businessFL_store, businessFL_Juice, businessFL_desserts, businessFL_events)
b = b %>% group_by(business_id) %>% filter (!duplicated(business_id)) # remove any duplicates
b %>% group_by(`Format Category`)%>%summarise(Count = n()) # group by category and count the number of businesses in each category



```

## Classified by restaurant Special of categories
Here, we divided it into Seafood, Vegetarian, Vegan, Salad, Gluten-Free
```{r}

# Function to categorize businesses by special category
categorize_special_business <- function(data, category, category_name) {
  categorized_data <- data.frame()
  for(i in 1:nrow(data)) {
    a = data[i,]
    if(sum(c(category) %in% unlist(strsplit(a$categories,", "))) > 0) {
      a = a %>% mutate(`Special Category` = category_name)
      categorized_data = rbind(categorized_data, a)
    }
  }
  return(categorized_data)
}

# Set up the list of special categories to search for
categories <- list(
  c("Seafood"),
  c("Vegan"),
  c("Vegetarian"),
  c("Salad"),
  c("Gluten-Free")
)

# Use the function to categorize businesses by special category
businessFL1 = businessFL
businessFL_special <- data.frame()

for(i in 1:length(categories)) {
  businessFL_special = rbind(
    businessFL_special,
    categorize_special_business(
      data = businessFL1, 
      category = categories[[i]], 
      category_name = toupper(categories[[i]])
    )
  )
}

# Group the businesses by special category and count the number of businesses in each category
businessFL_special = businessFL_special %>% group_by(business_id) %>% filter(!duplicated(business_id))
businessFL_special %>% group_by(`Special Category`) %>% summarise(Count = n())

```

## Deal with reviews
```{r}
reviews = read.csv("../../data/clean_data/reviews.csv",sep = '\t')
reviewsFL = reviews %>% filter(state == "FL")
```


```{r}
# select specific column
reviewsFL1 = reviewsFL%>%select(business_id,stars_x,review_count,categories,stars_y,useful,funny,cool,date)
```

```{r}
# Update date column to only include the first 10 characters
reviewsFL1 = reviewsFL1 %>% mutate(date=substr(date, 1,10))

# Create a data frame to hold reviews for businesses in the "Events" category
reviewsFL_events = data.frame()

# Loop through all reviews
for(i in 1:1082916){
  is_event = reviewsFL1[i,]
  
  # Check if the business associated with the review is in the "Events" category
  if(sum(c('Music Venues','Event Planning & Services','Arts & Entertainment','Nightlife','Venues & Event Spaces','') %in% unlist(strsplit(is_event$categories,", ")))>0){
    
    # If the business is in the "Events" category, add the review to the "Events" data frame
    reviewsFL_events = rbind(reviewsFL_events,is_event)
  }
}

# Create a new column indicating whether the business associated with the review is in the "Events" category
reviewsFL_events = reviewsFL1 %>% mutate(Events = ifelse(sum(c('Music Venues','Event Planning & Services','Arts & Entertainment','Nightlife','Venues & Event Spaces','') %in% unlist(strsplit(categories,", ")))>0, T, F))

# Filter the reviews data frame to only include reviews for businesses in the "Events" category
reviewsFL_events = reviewsFL_events %>% filter(Events==T)

```

```{r}
a1 = a %>%select(business_id, is_open, Business.Status, `Region Category`)
colnames(a1)[3] = "Business Status"
r1 = merge(a1, reviewsFL,by="business_id")
r1 = r1 %>% select(business_id, `Business Status`, `Region Category`, stars_y, date)
colnames(r1)[4]="Rating"
r1 = r1 %>% mutate(Year = substr(date,1,4))

```

# Open Region restaurants
```{r}
r1sum1 = r1 %>%select(-date)%>%filter(`Business Status` == "Open")
r1sum1 =r1sum1 %>%group_by(Year, `Region Category`)%>%summarise(Rating = mean(Rating),n=n())
r1sum1 = r1sum1[-1,]
r1sum1=r1sum1%>%filter(Year<2022)
write.csv(r1sum1, "Reviews_Rating_Region_Open.csv")
```


# Subset the datasets with different business status and different region category
```{r}
# Select relevant columns and filter for open businesses
r1sum1 = r1 %>%select(-date)%>%filter(`Business Status` == "Open")

# Group by year and region category, calculate average rating and count of reviews
r1sum1 =r1sum1 %>%group_by(Year, `Region Category`)%>%summarise(Rating = mean(Rating),n=n())

# Remove first row (no data for year 2012)
r1sum1 = r1sum1[-1,]

# Filter out data for year 2022 and later
r1sum1=r1sum1%>%filter(Year<2022)

# Write to CSV file
write.csv(r1sum1, "Reviews_Rating_Region_Open.csv")
```


# Close Region restaurants
```{r}
r1sum1 = r1 %>%select(-date)%>%filter(`Business Status` == "Close")
r1sum1 =r1sum1 %>%group_by(Year, `Region Category`)%>%summarise(Rating = mean(Rating),n=n())
r1sum1 = r1sum1[-1,]
r1sum1 = r1sum1 %>%filter(Year<2022)
write.csv(r1sum1, "Reviews_Rating_Region_Close.csv")
```

# Open Format restaurants
```{r}
a2 = b %>%select(business_id, is_open, Business.Status, `Format Category`)
colnames(a2)[3] = "Business Status"
r2 = merge(a2, reviewsFL,by="business_id")
r2 = r2 %>% select(business_id, `Business Status`, `Format Category`, stars_y, date)
colnames(r2)[4]="Rating"
r2 = r2 %>% mutate(Year = substr(date,1,4))
r1sum1 = r2 %>%select(-date)%>%filter(`Business Status` == "Open")
r1sum1 =r1sum1 %>%group_by(Year, `Format Category`)%>%summarise(Rating = mean(Rating),n=n())
r1sum1 = r1sum1[-1,]
write.csv(r1sum1, "Reviews_Rating_Format_Open.csv")
```
# Close Format restaurants
```{r}
r1sum1 = r2 %>%select(-date)%>%filter(`Business Status` == "Open")
r1sum1 =r1sum1 %>%group_by(Year, `Format Category`)%>%summarise(Rating = mean(Rating),n=n())
r1sum1 = r1sum1[-1,]
write.csv(r1sum1, "Reviews_Rating_Format_Close.csv")
```

# Open Special restaurants
```{r}
a3 = c %>%select(business_id, is_open, Business.Status, `Special Category`)
colnames(a3)[3] = "Business Status"
r3 = merge(a3, reviewsFL,by="business_id")
r3 = r3 %>% select(business_id, `Business Status`, `Special Category`, stars_y, date)
colnames(r3)[4]="Rating"
r3 = r3 %>% mutate(Year = substr(date,1,4))
r1sum1 = r3 %>%select(-date)%>%filter(`Business Status` == "Open")
r1sum1 =r1sum1 %>%group_by(Year, `Special Category`)%>%summarise(Rating = mean(Rating),n=n())
r1sum1 = r1sum1[-1,]
write.csv(r1sum1, "Reviews_Rating_Special_Open.csv")
```

# Close Special restaurants
```{r}
r1sum1 = r3 %>%select(-date)%>%filter(`Business Status` == "Close")
r1sum1 =r1sum1 %>%group_by(Year, `Special Category`)%>%summarise(Rating = mean(Rating),n=n())
r1sum1 = r1sum1[-1,]
write.csv(r1sum1, "Reviews_Rating_Special_Close.csv")
```



Combine the top photo data with the reviews data

```{r}
topfig = read.csv("../../data/clean_data/topfig.csv")
topfig = topfig[,-c(1,2)]
#topfig = topfig %>%filter(label=="food")
topfig = topfig %>% group_by(Year,Region.Category,Rating, n, business_id)%>%slice_max(photo_id)
top1 = topfig %>%group_by(Year, Region.Category,business_id,label)%>%summarise(n=n())
write.csv(topfig,"topfig1.csv")
```


```{r}
photo = read.csv("../../data/clean_data/photo.csv")
photo1 = photo%>%group_by(business_id)%>%summarise(n1=n())

topfig = read.csv("../../data/clean_data/topfig.csv")
topfig = topfig[,-c(1,2)]

# Merge topfig and photo1 dataframes based on the business_id column
topfig1 = merge(topfig,photo1,by="business_id")

# Group by Year and Region.Category and calculate the count
topfigsum = topfig1%>%group_by(Year,Region.Category)%>%summarise(n3 = n())

# For each unique combination of Year, Region.Category, Rating, and n, keep only the row with the maximum value of business_id
topfig11 = topfig1 %>% group_by(Year,Region.Category,Rating, n)%>%slice_max(business_id)

# Select business_id and photo_id columns from topfig1
photo1 = topfig1%>%select(business_id,photo_id)

# Merge topfig11 and photo1 dataframes based on the business_id column
topfig111 = merge(topfig11,photo1,by="business_id")

# Keep only the row with the maximum value of photo_id for each unique combination of Year, Region.Category, Rating, n, and business_id
topfig111 = topfig111 %>%select(business_id,Year,Region.Category,Rating,n,photo_id)
topfig111 = topfig111 %>%group_by(Year,Region.Category,Rating,n,business_id)%>%slice_max(photo_id)
topfig112=distinct(topfig111)

# Keep only the row with the maximum value of photo_id for each unique combination of Year, Region.Category, Rating, n, and business_id
topfig1 = topfig1 %>% group_by(Year,Region.Category,Rating, n, business_id)%>%slice_max(photo_id,business_id)
topfig1 = topfig1%>%group_by(Year,Region.Category,Rating, n, business_id)

# Count the number of instances of each unique combination of Year, Region.Category, business_id, and label
top1 = topfig %>%group_by(Year, Region.Category,business_id,label)%>%summarise(n=n())

# Merge topfig112 and name dataframes based on the business_id column
name = businessFL %>% select(business_id,name)
topfig112 = merge(topfig112,name,by="business_id",all.x = T)

# Remove duplicates from topfig112
topfig112=distinct(topfig112)

# Remove rows with missing values from topfig112
topfig112=topfig112%>%drop_na()

```

